---
title: "R Notebook"
output: html_notebook
---

# Setup

```{r setup, include=TRUE}
library(tidyverse)
library(tidytuesdayR)
library(rvest)
library(magrittr)
library(extrafont)
library(ggforce)
library(glue)
library(patchwork)
library(remotes)
library(jkmisc)
library(ggtext)
library(zoo)
```

```{r my.theme, include=TRUE}
my.theme <- theme(
  plot.title = element_text(family="Gill Sans MT", color="black", face="bold", size=24, hjust=0), 
  plot.subtitle = element_text(family="Gill Sans MT", color="black", size=12, hjust=0), 
  axis.title = element_text(family="Gill Sans MT", color="black", face="bold", size=12), 
  axis.text = element_text(family="Gill Sans MT", color="black", size=14), 
  axis.ticks = element_blank(),
  #axis.ticks.y = element_blank(), 
  plot.caption = element_text(family="Gill Sans MT", color="black", size=12), 
    panel.background =  element_blank(), 
  panel.grid.major = element_line(colour = "grey90", size = 0.5),
  panel.grid.minor = element_line(colour = "grey93", size = 0.5),
  #panel.background =  element_rect(fill = "#F7F7F7", colour = NA), 
  #legend.background = element_rect(fill = "#F7F7F7", colour = NA),
  panel.border = element_blank(), 
  #panel.border = element_rect(colour = "black", size = 0.5, fill=NA, linetype = 1),
  legend.title=element_blank(), 
  legend.text = element_text(family="Gill Sans MT", color="black", size=12, hjust=0),
  legend.position = 'top',
  strip.text = element_text(family="Gill Sans MT", color="black", size=14))
```

```{r read_data, include=TRUE}
tuesdata <- tidytuesdayR::tt_load(2020, week = 2)
#fire_nrt <- tuesdata$fire_nrt_M6_94771
rainfall <- tuesdata$rainfall
temperature <- tuesdata$temperature
```

```{r bind_df, include=FALSE}
df <- bind_rows(
  tuesdata$IDCJAC0009_009151_1800_Data,
  tuesdata$IDCJAC0009_023011_1800_Data,
  tuesdata$IDCJAC0009_040383_1800_Data,
  tuesdata$IDCJAC0009_040913_1800_Data,
  tuesdata$IDCJAC0009_066062_1800_Data,
  tuesdata$IDCJAC0009_070351_1800_Data,
  tuesdata$IDCJAC0009_086232_1800_Data)
```

```{r df_summary, include=TRUE}
df %>% 
  summary()
```

# EDA  

```{r annual_rainfall, include=TRUE}
annual_rainfall <- rainfall %>% 
  group_by(year, month) %>% 
  summarize(avg_monthly_rainfall = mean(rainfall, na.rm = TRUE)) %>% 
  group_by(year) %>% 
  summarize(annual_rainfall = sum(avg_monthly_rainfall, na.rm = TRUE))
```

```{r plot_rainfall, fig.height=6.2}
(plot_rainfall <- annual_rainfall %>% 
  filter(year >= 1900 & year < 2020) %>% 
  mutate(ma10 = zoo::rollapply(annual_rainfall, width=10, FUN=mean, align='right', partial = TRUE)) %>% 
  ggplot(.,
         aes(x=year,
             y=annual_rainfall)) + 
  geom_line(color = 'cornflowerblue') + 
  geom_point(color = 'cornflowerblue') + 
  geom_line(aes(y=ma10), 
            linetype='longdash', color='gray30') + 
  geom_mark_circle(aes(filter = year == 2019, 
                       label = glue("Total Rainfall: {round(annual_rainfall,1)} mm"), 
                       description = "Annual rainfall at a 60 year low"), 
                   label.family = c("Gill Sans MT", "Times New Roman"), label.buffer = unit(6, "cm"), label.fontsize = 10) + 
  scale_x_continuous(limits = c(1900, 2020), breaks = seq(1900, 2020, 20)) +
  labs(x='',
       y='') + 
  scale_y_continuous(limits = c(10, 50), breaks = seq(10, 50, 10), label = label_number_si(unit = "mm")) +
  my.theme)

```

For the temperature, I am finding that I can't exactly replicate the code from `jkaupp`. It could be that this is not taking into account a monthly average first (taking an average of averages like in rainfall). The most likely reason is that he probably pulled in data that included all of 2019, instead of just downloading through the `tidytuesday` package.  

```{r months_represented, include=TRUE}
rainfall %>% 
  filter(year == 2019) %>% 
  group_by(month) %>% 
  summarise(n_distinct(city_name))

temperature %>% 
  filter(temp_type == 'max') %>% 
  mutate(year = year(date),
         month = month(date)) %>% 
  filter(year == 2019) %>% 
  group_by(month) %>% 
  summarise(n_distinct(city_name))
```

```{r make_annual_temperature, include=TRUE}
annual_temperature <- temperature %>% 
  filter(temp_type == 'max') %>% 
  mutate(year = year(date)) %>% 
  group_by(year) %>% 
  summarise(avg_temperature = mean(temperature, na.rm = TRUE)) %>% 
  filter(year < 2020)
```

```{r plot_temperature, include=TRUE, fig.height=6.2}
(plot_temperature <- annual_temperature %>% 
  mutate(ma10 = zoo::rollapply(avg_temperature, width=10, FUN=mean, align='right', partial = TRUE)) %>% 
  ggplot(.,
         aes(x=year,
             y=avg_temperature)) + 
  geom_line(color = 'firebrick3') + 
  geom_point(color = 'firebrick3') + 
  geom_line(aes(y=ma10),
            linetype='longdash', color='gray30') + 
  geom_mark_circle(aes(filter = year == 2019, 
                       label = glue("Avg Max Temp: {round(avg_temperature, 1)}°C"),
                       description = "Data through May 2019"), 
                   label.family = c("Gill Sans MT", "Times New Roman"), label.buffer = unit(6, "cm"), label.fontsize = 10) + 
  labs(x='',
       y='') + 
  scale_x_continuous(limits = c(1900, 2020), breaks = seq(1900, 2020, 20)) +
  scale_y_continuous(limits = c(10, 30), breaks = seq(10, 30, 10), label = label_number_si(unit = "° C", sep = ""), position = "right") +
  my.theme)
```

To do this advanced formatting, I need to use `labs()` instead of `ggtitle()` in my plots. 

```{r plot_rain_temp, include=TRUE, fig.height=8}
(plot_rain_temp <- plot_rainfall + 
  plot_temperature + 
  plot_annotation(title = "Historic Lows in Rainfall and Records High Temperatures a Factor in Australia's Devastating Wildfires",
                  subtitle = glue('{highlight_text("Total rainfall", "cornflowerblue", "b")} and {highlight_text("average maximum temperature", "firebrick3", "b")} across Australian cities from 1900 to 2019<br>10-year moving average shown as dotted line'),
                  caption = 'Tidy Tuesday Week 2 (2020)<br>**Source**: Australian Bureau of Meterology',
                  theme = theme_jk(markdown = TRUE,
                                   plot_title_family = 'Gill Sans MT',
                                   subtitle_family = 'Gill Sans MT',
                                   caption_family = 'Gill Sans MT')))
```

# Saving graphics  

```{r save_graphics, include=TRUE}
ggsave("graphics/plot_rain_temp_1900_2019.png", plot_rain_temp,
       width = 12, height = 8, dpi = 300)
```

