---
title: "R Notebook"
output: html_notebook
---

```{r setup, include=TRUE}
library(tidyverse)
library(tidytuesdayR)
library(rvest)
library(magrittr)
library(extrafont)
library()
```

```{r read_data, include=TRUE}
tuesdata <- tidytuesdayR::tt_load(2020, week = 38)
kids <- tuesdata$kids
```

# EDA  

```{r glimpse, include=TRUE}
kids %>% 
  glimpse()
```

```{r summary, include=TRUE}
kids %>% 
  summary()
```

```{r distinct_variable, include=TRUE}
kids %>% 
  distinct(variable)
```

**PK12ed**: Public spending on elementary and secondary education by state and year, in $1,000s. Data are from the Census Bureau's annual State and Local Government Finance Survey, expenditure variable E12 less private revenues from School Lunch (A09), School Tuition (A10), and Other Elementary-Secondary Education (A12). Data include public prekindergarten spending by school districts, including federal Head Start grants to school districts.

```{r PK12ed_hist, include=TRUE, fig.height=6.2}
kids %>% 
  filter(variable == 'PK12ed') %>% 
  ggplot(.) + 
  geom_histogram(aes(x=inf_adj_perchild)) + 
  ggtitle('Distribution of PK12ed')
```

```{r url_scrape, include=TRUE}
url <- 'https://en.wikipedia.org/wiki/List_of_United_States_governors'

governors <- 
  url %>% 
  read_html() %>% 
  html_nodes(xpath='//*[@id="mw-content-text"]/div[1]/table[1]') %>% 
  html_table(fill = TRUE)

governors <- data.frame(governors) %>% 
  select(1, 3, 5) %>% 
  slice(-1) %>% 
  rename(state = Democratic..See.Tfd...24.....Republican..See.Tfd...26.,
         name = Democratic..See.Tfd...24.....Republican..See.Tfd...26..2,
         party = Democratic..See.Tfd...24.....Republican..See.Tfd...26..4)

dc_governor <- 
  url %>% 
  read_html() %>% 
  html_nodes(xpath='//*[@id="mw-content-text"]/div[1]/table[3]') %>% 
  html_table(fill = TRUE)

dc_governor <- data.frame(dc_governor) %>% 
  rename(state = Democratic..See.Tfd...1.,
         name = Democratic..See.Tfd...1..2,
         party = Democratic..See.Tfd...1..4)

governors <- bind_rows(governors, dc_governor)
```

```{r recoding, include=TRUE}
governors %<>% 
  mutate(party = recode(party,
                        'Democratic–Farmer–Labor' = 'Democratic',
                        'Republican[note 1]' = 'Republican'))
```

```{r make_kids_2016, include=TRUE}
kids_2016 <- kids %>% 
  filter(variable == 'PK12ed',
         year == 2016) %>% 
  left_join(x=.,
            y=governors,
            by='state')
```

```{r party_cols, include=TRUE}
party_cols <- c('Republican' = 'red1',
                'Democratic' = 'blue1')
```

```{r my.theme, include=TRUE}
my.theme <- theme(
  plot.title = element_text(family="Gill Sans MT", color="black", face="bold", size=16, hjust=0), 
  plot.subtitle = element_text(family="Gill Sans MT", color="black", size=14, hjust=0), 
  axis.title = element_text(family="Gill Sans MT", color="black", face="bold", size=12), 
  axis.text = element_text(family="Gill Sans MT", color="black", size=11), 
  axis.ticks.y = element_blank(), 
  plot.caption = element_text(family="Gill Sans MT", color="black", size=12), 
  plot.background = element_rect(fill = "#F7F7F7", colour = NA),
  panel.background =  element_rect(fill = "#F7F7F7", colour = NA), 
  legend.background = element_rect(fill = "#F7F7F7", colour = NA), 
  panel.grid.major = element_blank(),
  panel.grid.minor = element_blank(),
  panel.border = element_blank(), 
  #panel.border = element_rect(colour = "black", size = 0.5, fill=NA, linetype = 1),
  legend.title=element_blank(), 
  legend.text = element_text(family="Gill Sans MT", color="black", size=12, hjust=0),
  legend.position = 'top',
  strip.text = element_text(family="Gill Sans MT", color="black", size=14))
```

```{r top_10_viz, include=TRUE, fig.height=6.2}
(top_10_viz <- kids_2016 %>% 
  arrange(-inf_adj_perchild) %>% 
  slice(1:10) %>% 
  ggplot(.) + 
  geom_col(aes(x=inf_adj_perchild,
               y=reorder(state, inf_adj_perchild),
               fill=party)) + 
  geom_text(aes(label=round(inf_adj_perchild, 1),
                x=inf_adj_perchild, 
                y=reorder(state, inf_adj_perchild)), 
            vjust=0.5, 
            size = 4, 
            position = position_stack(vjust = 0.5),
            color='grey98',
            fontface='bold') +
  ggtitle('Per child expenditure on public education by state in 2016 - Top 10 states',
          subtitle = "Party represents the Governor's party from Wikipedia") + 
  labs(x='Per child expenditure (inflation adjusted, 1000s US$)',
       y='',
       fill='',
       caption='\nTidy Tuesday: Week 38 (2020)\nData source: Urban Institute\nGovernor party not necessarily matches the correct party in 2016') + 
  scale_fill_manual(values = party_cols) + 
  my.theme)
```

```{r bottom_10_viz, include=TRUE, fig.height=6.2}
(bottom_10_viz <- kids_2016 %>% 
  arrange(inf_adj_perchild) %>% 
  slice(1:10) %>% 
  ggplot(.) + 
  geom_col(aes(x=inf_adj_perchild,
               y=reorder(state, inf_adj_perchild),
               fill=party)) + 
  geom_text(aes(label=round(inf_adj_perchild, 1),
                x=inf_adj_perchild, 
                y=reorder(state, inf_adj_perchild)), 
            vjust=0.5, 
            size = 4, 
            position = position_stack(vjust = 0.5),
            color='grey98',
            fontface='bold') +
  ggtitle('Per child expenditure on public education by state in 2016 - Bottom 10 states',
          subtitle = "Party represents the Governor's party from Wikipedia") + 
  labs(x='Per child expenditure (inflation adjusted, 1000s US$)',
       y='',
       fill='',
       caption='\nTidy Tuesday: Week 38 (2020)\nData source: Urban Institute\nGovernor party not necessarily matches the correct party in 2016') + 
  scale_fill_manual(values = party_cols) + 
  my.theme)
```

# Saving graphics  

```{r save_graphics, include=TRUE}
ggsave("graphics/top_10_viz.png", top_10_viz,
       width = 10, height = 8, dpi = 300)

ggsave("graphics/bottom_10_viz.png", bottom_10_viz,
       width = 10, height = 8, dpi = 300)
```
